---
title: "Assign. 1 STA 445"
author: "Myleen Maldonado"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```
## Directions:
This assignment covers chapter 5.  Please show all work in this document and knit your final draft into a pdf.  This is assignment is about statistical models, which will be helpful if you plan on taking STA 570, STA 371, or STA 571. 

## \color{black} Problem 1: Two Sample t-test

##\color{blue} 

a. Load the `iris` dataset.
```{r}
slice_sample(iris)
```


b. Create a subset of the data that just contains rows  for the two species setosa and versicolor using filter.  Use slice_sample to print out 20 random rows of the dataset.


```{r}
subset.iris <- iris %>%
  filter(Species=="setosa"| Species=="versicolor")
 slice_sample(subset.iris, n = 20)
```


c. Create a box plot of the petal lengths for these two species using ggplot. Does it look like the mean petal length varies by species?
```{r}
ggplot(subset.iris, aes(x=Species, y=Sepal.Length)) + geom_boxplot()

```


d. Do a two sample t-test using t.test to determine formally if the petal lengths differ. Note:  The book uses the tidy function in the broom package to make the output "nice".  I hate it!  Please don't use tidy.
```{r}
t.test(data=subset.iris, Petal.Length~Species, conf.level=0.9)
```

d. What is the p-value for the test? What do you conclude?
```{r}
# We conclude that because p-value < 2.2e-16 , then we can conclude that they are similar.
```
e. Give a 95% confidence interval for the difference in the mean petal lengths.  
```{r}
t.test(data=subset.iris, Petal.Length~Species, conf.level=0.95)

```

f. Give a 99% confidence interval for the difference in mean petal lengths.  (Hint: type ?t.test.  See that you can change the confidence level using the option conf.level)
```{r}
t.test(data=subset.iris, Petal.Length~Species, conf.level=0.99)

```
g. What is the mean petal length for setosa?
```{r}
# 1.462
```
h. What is the mean petal length for versicolor?
```{r}
# 4.260 

```

## Problem 2: ANOVA 

Use the iris data with all three species.

a. Create a box plot of the petal lengths for all three species using ggplot.Does it look like there are differences in the mean petal lengths?
```{r}
ggplot(iris, aes(x=Species, y=Sepal.Length)) + geom_boxplot()

#yes it looks like there are differences between petal lengths
```
b. Create a linear model where sepal length is modeled by species. Give it
an appropriate name.
```{r}

sepal.lgth <- lm(data=iris, Sepal.Length~Species-1)

sepal.lgth
```
c. Type anova(your model name) in a code chunk.
```{r}
anova(sepal.lgth)
```

d. What is the p-value for the test?  What do you conclude.
```{r}
# it ends up being smaller than 2.2e-16 so that means that they are similiar.
```

e. Type summary(your model name) in a code chunk.
```{r}
summary(sepal.lgth)
```

f. What is the mean petal length for the species setosa?
```{r}
#  the mean petal length for species setosa is:
#  1.6880
```

g. What is the mean petal length for the species versicolor?
```{r}
#  the mean petal length for species versicolor is:
#  1.582  

sepal.lgth <- lm(data=iris, Sepal.Length~Species)

sepal.lgth

```

## Problem 3: Regression 

Can we describe the relationship between petal length and petal width?

a. Create a scatterplot with petal length on the y-axis and petal width on the x-axis using ggplot.
```{r}
ggplot(iris, aes(x=Petal.Width, y=Petal.Length)) + 
  geom_point()

```
b. Create a linear model to model petal length with petal width (length is the response variable and width is the explanatory variable) using lm.
```{r}
petal.lm <- lm(data=iris, Petal.Length~Petal.Width * Species)

petal.lm
```
c. What is the estimate of the slope parameter?
```{r}
petal.lm
  #  1.3276  

```
d. What is the estimate of the intercept parameter?
```{r}
summary(petal.lm)
  # 0.1309
```
e. Use summary() to get additional information.
```{r}
summary(petal.lm)
```

## Problem 4: Modeling Trees

Using the `trees` data frame that comes pre-installed in R, follow the steps below to fit the regression model that uses the tree `Height` to explain the `Volume` of wood harvested from the tree.

a. Create a scatterplot of the data using ggplot.
```{r}
ggplot(trees, aes(x=Volume, y=Height)) + 
  geom_point()


```

b. Fit a `lm` model using the command `model <- lm(Volume ~ Height, data=trees)`.
```{r}
model <- lm(Volume ~ Height, data=trees)
model
```
c. Print out the table of coefficients with estimate names, estimated value, standard error, and upper and lower 95% confidence intervals.
```{r}
summary(model)


```
d.Add the model fitted values to the `trees` data frame along with the regression model confidence intervals.  Note: the book does this in a super convoluted way.  Don't follow the model in the book.  Instead try cbind.

```{r}
trees.pred <- cbind(trees, predict(model, interval="confidence"))

head(trees.pred)
```
e. Graph the data and fitted regression line and uncertainty ribbon.
```{r}
ggplot(data = trees.pred, aes(x=Volume, y=Height)) +
  geom_point(size=3) +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill = "skyblue")
```
f. Add the R-squared value as an annotation to the graph using annotate.
```{r}
ggplot(data = trees.pred, aes(x=Volume, y=Height)) +
  geom_point(size=3) +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill = "skyblue") 
 # annotate("r-squared value is:" )



```