---
title: "STA_445_Assignment_6"
author: "Myleen"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(mosaicData)
```

## Problem 1
Convert the following to date or date/time objects.

a. September 13, 2010.
```{r}
p1a <- mdy("September 13, 2010")
p1a
class(p1a)
```

b. Sept 13, 2010.
```{r}
p1b <- mdy("September 13, 2010")
p1b
class(p1b)
```


c. Sep 13, 2010.
```{r}
p1c <- mdy("September 13, 2010")
p1c
class(p1c)
```

d. S 13, 2010. Comment on the month abbreviation needs.
```{r}
#incorrect
#p1d <- mdy("S 13, 2010")
#p1d
#class(p1d)

#corrected
p1d <- mdy("Sep 13, 2010")
p1d
class(p1d)

#instead of S it needs to be Sep
```

e. 07-Dec-1941.
```{r}
p1e <- dmy("07-Dec-1941.")
p1e
class(p1e)
```

f. 1-5-1998. Comment on why you might be wrong.
```{r}
p1f <- mdy("1-5-1998.")
p1f
class(p1f)
#I might be wrong because both month/day should be double digit and there is overlap between 1 and 12 for both month/day to decide which is which. Therefore, it is best to use abbreviations of letters when naming the months
```

g. 21-5-1998. Comment on why you know you are correct.
```{r}
p1g <- dmy("21-5-1998.")
p1g
class(p1g)

# I know I am right because there is no month that is known as 21, therefore I know that it this goes day/month/year
```

h. 2020-May-5 10:30 am
```{r}
p1h <- ymd_hm("2020-May-5 10:30 am")
p1h
class(p1h)
```


i. 2020-May-5 10:30 am PDT (ex Seattle)
```{r}
p1i<- ymd_hm("2020-May-5 10:30 am",tz= "America/Los_Angeles")
p1i
class(p1i)
```


j. 2020-May-5 10:30 am AST (ex Puerto Rico)
```{r}
p1j<- ymd_hm("2020-May-5 10:30 am",tz= "America/Puerto_Rico")
p1j
class(p1j)
```

### Problem 2

Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following:

a. Calculate the date of your 64th birthday.
```{r}
bday <- mdy("09/30/2004", tz="US/Arizona")
bday64 <- bday + years(64)
bday64
```

b. Calculate your current age (in years). 
```{r}
bday %--% now()
as.period(bday %--% now())
as.duration(bday %--% now())
```
c. Using your result in part (b), calculate the date of your next birthday.
```{r}
nextBday <- bday + years(20)
nextBday
```

d. The number of _days_ until your next birthday.
```{r}
daysUntilBday <- as.period(now() %--% nextBday, unit="day" )
daysUntilBday

```
e. The number of _months_ and _days_ until your next birthday.
```{r}
monDayUntilBday <- as.period(now()%--%nextBday)

monDayUntilBday
```

## Problem 3

Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 
```{r}
phonecall <- mdy_hm("May 8, 2015 3:00pm", tz="US/Arizona")
phonecall

#with_tz(phonecall, tz="UTC")
with_tz(phonecall, tz="Pacific/Auckland")


```

## Problem 4

It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 

```{r}
#view(Births78)
#load("Births78")
births78 <- select(Births78, "date", "births")
head(births78)

```

b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?

```{r}
ggplot(births78, aes(x=date, y=births))+
  geom_point()

```

c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.
```{r}
new<-births78%>%
 mutate(dow = wday(births78$date, label=TRUE))
head(new)
```

d. Plot the data with the point color being determined by the day of the week variable.

```{r}
ggplot(new, aes(x=date, y=births, color=dow)) +
  geom_point()
```








